
RUNNING OUR TESTS*

```{r}
#libraries
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(foreign)
library(tidyverse)
library(epitools)
library(lmtest)
library(tidyverse)
library(tidyr)
library(broom)
library(epiR)
library(dplyr)
library(dagitty)
library(ggdag)
library(ggplot2)
library(dplyr)
install.packages("readxl")
library("readxl")
```

```{r}
# xlsx files
finalproject <- read_excel("NISPUF20_dataset.xlsx")
```

What are we doing?
The goal is to study the association between parity and breast feeding duration
Hypothesis: First time mothers have shorter breast feeding duration compared to mothers of 2 or more children

Steps:
- Identify the exposure and outcome variables
     - Exposure: FRSTBRN (NO/YES) binary variable
     - Outcome: BF_ENDR06 (NA - numbers) in days
          - Splitting into 4 categories:
               - less than 1 month (30 days)
               - 1 to less than 6 months (30 - 180 days)
               - 6 months to less than 12 months (180 - 360 days)
               - more than 12 months (360 - inf days)
          - How to handle N/A?
               - Omit
     - Potential Confounders: 
              DESCRIPTION     |  NAME      |     DATA FACTOR
        ----------------------|------------|---------------------------------------------------------------
          maternal age        | M_AGEGRP2  |  2 factors <=29 years and >= 30 years
          marital status      | MARITAL2   |  2 factors MARRIED and "OTHERS"
          maternal education  | EDUC1      |  4 factors (<12 yrs, 12 years, > 12 yrs-non-grad, college grad)
          child race          | RACEETHK   |  4 factors (HISPANIC, BLACK, OTHER + MULTIPLE RACE, WHITE)
          poverty status      | INCPOV1    |  4 factors (BELOW, ABOVE <= 75k, ABOVE > 75k, UNKNOWN)
          WIC Benefits        | CWIC_01    |  5 factors (NO, YES, REFUSED, DON'T KNOW, NEVER HEARD OF WIC)

- Analysis methods
      - Create the 4 categories for breastfeeding duration
      - Descriptive statistical analysis (mean, median, and mode to get central value and 
      find the measure of our data spread)
          - The distribution of breastfeeding duration
          - Summarize breastfeeding by parity to get the group means and standard deviations
          - Create a visualization that shows breastfeeding duration by parity
          - Create a visualization to check the normality of our outcome variable
      - Use logistic regression 
          - the model estimates probability of breastfeeding for specific number of days (the 4 categories)
          - compare models (looking at the estimate and p-value for each variable) to identity the statistically 
          significant confounders
      - Controling for confounding
      - Sample matching

```{r}

#Omitting NA from the outcome variable 
finalproject <- na.omit(finalproject)
```

```{r}
#Switching variables to categories 
finalproject$FRSTBRN <- factor(finalproject$FRSTBRN)
finalproject$MARITAL2 <- factor(finalproject$MARITAL2)
```

```{r}
summary(finalproject)
```

### Descriptive Analysis

```{r}
# # The distribution of breastfeeding duration
# hist(finalproject$BF_ENDR06)
# 
# #Creating a visualization that shows breastfeeding duration by parity
# ggplot(finalproject, aes(group=FRSTBRN, x=FRSTBRN, y = BF_ENDR06)) + 
#   geom_boxplot() +
#   theme_minimal(base_size = 10) 
# 
# # or without a boxplot
# ggplot(finalproject, aes(x=FRSTBRN, y=BF_ENDR06)) +
#     geom_point(shape=1) +
#     labs(y = "Average Breastfeeding Duration ",
#          x = "First Born", title = "Breastfeeding Duration vs First Born")
# 
# #Creating a visualization to check the normality of our outcome variable (breastfeeding duration)
# finalproject %>%
#   ggplot(aes(x=BF_ENDR06)) + geom_histogram(colour="black", fill="white") + facet_wrap(~FRSTBRN)+
#     labs(title = "A histogram plot for Breastfeeding Duration (in days) per parity")
# 
# #or
# finalproject %>%
#   ggplot(aes(sample=BF_ENDR06)) + stat_qq(colour="black", fill="white") + facet_wrap(~FRSTBRN)+
#     labs(title = "A quantile-quantile plot for Breastfeeding Duration (in days) per parity")
```
- Summarize breastfeeding by parity to get the group means and standard deviations
```{r}
finalproject%>% 
  group_by(FRSTBRN) %>% 
  summarize(mean_sdp = mean(BF_ENDR06, na.rm=TRUE),sd_sdp=sd(BF_ENDR06, na.rm=TRUE))
```

## making interaction terms binary
```{r}

unique(finalproject$CWIC_01)
finalproject <- finalproject %>%
  mutate(wic = case_when(CWIC_01 == "YES" ~ 1, 
                        CWIC_01 == "NO" | CWIC_01 == "NEVER HEARD OF WIC" ~ 0,
                        TRUE ~ NA_real_))

unique(finalproject$INCPOV1)
finalproject <- finalproject %>%
  mutate(ses = case_when(INCPOV1 == "ABOVE POVERTY, > $75K" | INCPOV1 == "ABOVE POVERTY, <= $75K" ~ 1, 
                         INCPOV1 == "UNKNOWN" ~ NA_real_,
                         TRUE ~ 0))

unique(finalproject$EDUC1)
finalproject <- finalproject %>%
  mutate(ed = if_else(EDUC1 == "COLLEGE GRAD",1,0))

finalproject <- finalproject %>%
  dplyr::select(c(-INCPOV1, -CWIC_01, -EDUC1)) %>%
  rename(bf = BF_ENDR06, maternal_age = M_AGEGRP2, marital_status = MARITAL2, race_eth = RACEETHK, rent_own = RENT_OWN, state = STATE, first_born = FRSTBRN, child_age = AGEGRP)
```


### running the model using LINEAR REGRESSION 
  (with a continuous outcome)
```{r}
library(knitr)
library(foreign)
library(tidyverse)
library(epitools)
library(lmtest)
library(epiR)
library(car)

#Running linear regression model
model_0 <- lm(bf ~ first_born, data=finalproject)
summary(model_0)

model_1 <- lm(bf ~ first_born + ed, data = finalproject)
summary(model_1)

model_2 <- lm(bf ~ first_born + maternal_age, data = finalproject)
summary(model_2)

model_sat <- lm(bf ~ first_born + maternal_age + ed + ses + wic, 
              data = finalproject)
summary(model_sat)


##comparing the restricted model to the saturated model
anova(model_0, model_sat)
```
### Use logistic regression 
- the model estimates probability of breastfeeding for specific number of days (the 4 categories)
```{r}
# # Creating the 4 outcome categories
 finalproject_cat  <-
   finalproject  %>%
   dplyr::mutate(bf_cat = case_when(
     bf < 30 ~ "< 1 month",
     bf < 180 ~ "1 - 6 months",
     bf < 360 ~ "6 - 12 months",
     bf >= 360 ~ "> 12 months"))

#Omitting NA from the outcome variable 
finalproject_cat <- na.omit(finalproject_cat)

#Switching variables to categories 
finalproject_cat$bf_cat <- factor(finalproject_cat$bf_cat)
finalproject_cat$marital_status <- factor(finalproject_cat$first_born)
finalproject_cat$marital_status <- factor(finalproject_cat$marital_status)
```

```{R}
#Running logistic regression model
model_00 <- glm(bf_cat ~ first_born,
                     family="binomial", data=finalproject_cat)
summary(model_0)

model_01 <- glm(bf_cat ~ first_born + ed, family = "binomial", data = finalproject_cat)
summary(model_1)

model_02 <- glm(bf_cat ~ first_born + maternal_age, family = "binomial", data = finalproject_cat)
summary(model_2)

model_0sat <- glm(bf_cat ~ first_born + ed + maternal_age + ed:ses + ed:maternal_age, family = "binomial", data = finalproject_cat)
summary(model_0sat)

##comparing the restricted model to the saturated model
lrtest(model_00, model_01, model_02, model_0sat)


#assuming education is not a confounder
model_0sat_alt <- glm(bf_cat ~ first_born + maternal_age + ed:maternal_age + ses:maternal_age, family = "binomial", data = finalproject_cat)
summary(model_0sat_alt)

#comparing the saturated model and the alternative saturated model (education not a confounder)
lrtest(model_0sat, model_0sat_alt)

### comparing model with WIC status to SES model ** THIS DOESN'T WORK BECAUSE WIC IS NOT CONNECTED TO ANY OF OUR PREDICTORS DIRECTLY, THEREFOR IT SHOULDNT BE IN OUR MODEL.
# model_05 <- glm(bf_cat ~ first_born + ed + maternal_age + ed:ses, wic:ses + maternal_age:ses, family = "binomial", data = finalproject_cat)
# model_06 <- glm(bf_cat ~ first_born + ed + maternal_age + ed:ses + ses:maternal_age, family = "binomial", data = finalproject_cat)
# 
# lrtest(model_05, model_06)


## Running miltimodel regression model for 4 categories
# library(nnet)
# model_2 <- multinom(BF_ENDR06_CAT ~ FRSTBRN,
#                      family="binomial", data=finalproject_cat)
# 
# summary(model_2)
```
- compare models (looking at the estimate and p-value for each variable) to identity the statistically 
significant confounders

















##DAG##
  where 
  FB = first-born
  BF = breastfeeding duration
  AGE = maternal age
  ED = maternal education level
  SES = maternal poverty level
  WIC = participation in WIC program
  RACE = race/ethnicity of child
  RENT = rent/own status of mother
  STATE = state of residence
  WED = marital status
  SUP = unmeasured maternal support
  
```{r}

bf_dag <- dagitty('dag{
    FB [pos="0.5,1"]
    BF [pos="2,1"]
    AGE [pos="0.2,1.4"] 
    ED [pos="0,0.8"]
    SES [pos=".2,.4"]
    RACE [pos ="1, 0.45"]
    RENT [pos = "1.5, 0.6"]
    STATE[pos = "1.5, 1.5"]
    SUP[pos = "1.25, 1.05"]
    FB -> BF
    FB -> AGE
    AGE -> FB
    AGE -> BF
    AGE -> ED
    ED -> BF
    ED -> AGE
    SES -> BF
    SES -> ED
    SES -> AGE
    RACE -> SES
    RACE -> RENT
    RACE -> ED
    RENT -> SES
    SES -> RENT
    AGE -> RENT
    STATE -> RENT
    STATE -> ED
    STATE -> ED
    STATE -> BF
    STATE -> SES
    SUP -> BF
    SUP -> ED
    SUP -> RENT
    SUP -> SES
}')
p <- ggdag::ggdag(bf_dag) + theme_dag()
p$layers[[3]]$mapping <- 
  aes(colour = c("Observed", "Unobserved")[as.numeric(name == "SUP") + 1])
p + scale_color_manual(values = c("black", "#cc2055")) +
  theme(legend.position = c(0.8, 0.8))
  
# REMOVED FROM THE DAGG
      # WIC [pos="1, 1.5"]
     # RENT -> WIC
      # SES -> WIC
     # WIC -> SES
     # WIC -> BF
    # RACE -> WIC
   #WED[pos = "0.5, 2.0"]
   # WED -> BF
   #   WED -> SUP
# ED -> FB

```


```{r}
# mode - median 
# logistic regression since outcome categorical -> estimates probablilt of breastfeeding for specific number od days 
# cross sectional study -OR or prevalence OR? attributable proportion? 
# or RR
# since exposure "Yes" to first born then duration lower - so lower RR (less than 1)
# risk -> not breastfeeding for a long time [1- P()]
# how to frame exposure & outcome -> try to account for confounding 

#categories (look for backup studies)
# category 1 -> less than two weeks 
# category 2 ->
# category 3 ->
# category 4 -> 

# discrete data for outcome so estimate treatment effect** - average days of breastfeeding 
# after controlling for confounding
# OR makes sense when we have categorical variable - SO OR IS NOT A GOOD INDICATOR
```


